---
title: "Website Behavior & Booking Funnel Optimization"
subtitle: "Analytics Objective 5 | Google Analytics 4 | Jan 1 – Apr 21, 2026"
author: "Kathleen Juarez"
date: today
format:
  revealjs:
    theme: simple
    css: ao5-presentation.css
    transition: fade
    slide-number: true
    progress: true
    embed-resources: true
    width: 1050
    height: 620
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(scales)
library(knitr)
library(kableExtra)

sessions_raw <- read_csv("A05_data_sessions.csv",
                         skip = 9, show_col_types = FALSE) |>
  rename(channel = 1, sessions = 2, total_users = 3) |>
  filter(!is.na(channel), channel != "")

events_raw <- read_csv("A05_data_ga4-events.csv",
                       skip = 9, show_col_types = FALSE) |>
  rename(event_name = 1, channel = 2, event_count = 3, total_users = 4) |>
  filter(!is.na(event_name), event_name != "")

extract_event <- function(event_nm, col_name) {
  events_raw |>
    filter(event_name == event_nm) |>
    select(channel, {{ col_name }} := total_users)
}

page_view        <- extract_event("page_view",            page_views)
form_start       <- extract_event("form_start",           form_starts)
booking_complete <- extract_event("booking_complete",     bookings)
schedule_appt    <- extract_event("Schedule_Appointment", sched_appt_users)
contact_us       <- extract_event("contact_us",           contact_us_users)

ga4_data <- sessions_raw |>
  select(channel, sessions) |>
  left_join(page_view,        by = "channel") |>
  left_join(form_start,       by = "channel") |>
  left_join(booking_complete, by = "channel") |>
  left_join(schedule_appt,    by = "channel") |>
  left_join(contact_us,       by = "channel") |>
  mutate(across(-channel, ~replace_na(., 0))) |>
  mutate(
    page_view_rate    = if_else(sessions > 0,    page_views  / sessions,   0),
    form_start_rate   = if_else(page_views > 0,  form_starts / page_views, 0),
    booking_rate      = if_else(form_starts > 0, bookings    / form_starts, 0),
    overall_conv_rate = if_else(sessions > 0,    bookings    / sessions,   0)
  )

totals <- ga4_data |>
  summarise(
    channel           = "TOTAL",
    sessions          = sum(sessions),
    page_views        = sum(page_views),
    form_starts       = sum(form_starts),
    bookings          = sum(bookings),
    sched_appt_users  = sum(sched_appt_users),
    contact_us_users  = sum(contact_us_users),
    overall_conv_rate = sum(bookings) / sum(sessions)
  )

site_avg <- totals$overall_conv_rate

theme_ao5 <- theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.y = element_line(color = "#edf0f7"),
    axis.ticks         = element_blank(),
    legend.position    = "none",
    plot.title         = element_text(face = "bold", size = 13, color = "#1a1a2e"),
    plot.subtitle      = element_text(size = 10, color = "#6b7280",
                                      margin = margin(b = 8))
  )
```

##  {.title-slide-custom}

:::: title-block
**Analytics Objective 5**

# Website Behavior &<br>Booking Funnel Optimization

Using GA4 event-level data to find where users drop off, what creates friction, and what drives bookings.

::: chip-row
[📅 Jan 1 – Apr 21, 2026]{.chip} [563 Sessions]{.chip} [6 Traffic Channels]{.chip} [R / tidyverse]{.chip}
:::
::::

------------------------------------------------------------------------

## Methodology {.section-slide}

**Descriptive Funnel Analysis** — GA4 event-level data, *total users* metric throughout

::::: columns
::: {.column width="52%"}
**Funnel Stages (Dependent Variable)**

| \#  | Stage            | Description        |
|-----|------------------|--------------------|
| 1   | Sessions         | Entry point        |
| 2   | Page Views       | Surface engagement |
| 3   | Form Start       | Booking initiated  |
| 4   | Booking Complete | Conversion         |
:::

::: {.column width="48%"}
**Traffic Channels (Independent Variable)**

-   Direct
-   Organic Search
-   Organic Social
-   Referral
-   Unassigned
-   Paid Search
:::
:::::

------------------------------------------------------------------------

## The Core Problem {.section-slide}

```{r}
#| label: fig-funnel-pres
#| fig-width: 8
#| fig-height: 4

funnel_long <- tibble(
  stage = factor(
    c("Sessions", "Page Views", "Form Start", "Booking\nComplete"),
    levels = c("Sessions", "Page Views", "Form Start", "Booking\nComplete")
  ),
  count     = c(totals$sessions, totals$page_views,
                totals$form_starts, totals$bookings),
  highlight = c(FALSE, FALSE, TRUE, FALSE)
)

ggplot(funnel_long, aes(x = stage, y = count, fill = highlight)) +
  geom_col(width = 0.52) +
  geom_text(aes(label = comma(count)),
            vjust = -0.45, size = 4.2, fontface = "bold", color = "#1a1a2e") +
  scale_fill_manual(values = c("FALSE" = "#7e82ad", "TRUE" = "#C0392B")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  labs(
    title    = "Most Users Exit Before Completing Any High-Intent Action",
    subtitle = paste0(
      "Page View Rate: ",       percent(totals$page_views  / totals$sessions,   accuracy = 0.1),
      "  |  Form Start Rate: ", percent(totals$form_starts / totals$page_views, accuracy = 0.1),
      "  |  Overall Conv.: ",   percent(site_avg, accuracy = 0.1)
    ),
    x = NULL, y = "Unique Users"
  ) +
  theme_ao5
```

::: callout-box
**91.9% drop-off** from page views to form starts — the critical friction point, consistent across all six channels
:::

------------------------------------------------------------------------

## Not All Traffic Is Equal {.section-slide}

```{r}
#| label: fig-channels-pres
#| fig-width: 8
#| fig-height: 4.2

ga4_data |>
  mutate(
    channel   = fct_reorder(channel, overall_conv_rate),
    above_avg = overall_conv_rate > site_avg
  ) |>
  ggplot(aes(x = channel, y = overall_conv_rate, fill = above_avg)) +
  geom_col(width = 0.52) +
  geom_text(aes(label = percent(overall_conv_rate, accuracy = 0.1)),
            hjust = -0.15, size = 3.8, fontface = "bold", color = "#1a1a2e") +
  geom_hline(yintercept = site_avg, linetype = "dashed",
             color = "#9ca3af", linewidth = 0.55) +
  annotate("text", x = 0.65, y = site_avg + 0.0025,
           label = paste0("Site avg: ", percent(site_avg, accuracy = 0.1)),
           size = 3.2, color = "#9ca3af", hjust = 0) +
  coord_flip() +
  scale_fill_manual(values = c("FALSE" = "#b0bcd4", "TRUE" = "#7e82ad")) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.28))) +
  labs(
    title    = "Organic Search Is the Only Tracked Channel Above Site Average",
    subtitle = "Conv. Rate = booking_complete users ÷ sessions",
    x = NULL, y = NULL
  ) +
  theme_ao5 +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(color = "#edf0f7"))
```

------------------------------------------------------------------------

## Funnel Performance by Channel {.section-slide}

```{r}
#| label: tbl-funnel-pres

ga4_data |>
  arrange(desc(overall_conv_rate)) |>
  bind_rows(totals) |>
  mutate(overall_conv_rate = percent(overall_conv_rate, accuracy = 0.1)) |>
  select(
    `Traffic Channel` = channel,
    Sessions          = sessions,
    `Page Views`      = page_views,
    `Contact Us`      = contact_us_users,
    `Form Starts`     = form_starts,
    `Sched. Appts`    = sched_appt_users,
    `Conv. Rate`      = overall_conv_rate
  ) |>
  kbl(align = c("l","c","c","c","c","c","c")) |>
  kable_styling(bootstrap_options = c("condensed", "bordered"),
                full_width = TRUE, font_size = 13) |>
  row_spec(0, bold = TRUE, background = "#7e82ad", color = "white") |>
  row_spec(seq(1, nrow(ga4_data), by = 2), background = "white",   color = "#333") |>
  row_spec(seq(2, nrow(ga4_data), by = 2), background = "#f3f5f9", color = "#333") |>
  row_spec(nrow(ga4_data) + 1, bold = TRUE,
           background = "#e8edf6", color = "#1a1a2e") |>
  footnote(
    general = "Contact Us and Sched. Appts are high-intent signals (link clicks), not sequential funnel gates.",
    general_title = "", footnote_as_chunk = TRUE
  )
```

------------------------------------------------------------------------

## Subgroup Validation {.section-slide}

Drop-off from page views → form starts, confirmed across all channels

```{r}
#| label: tbl-dropoff-pres

ga4_data |>
  mutate(dropoff = 1 - form_start_rate) |>
  arrange(dropoff) |>
  select(
    `Traffic Channel` = channel,
    `Page Views`      = page_views,
    `Form Starts`     = form_starts,
    `Form Start Rate` = form_start_rate,
    `Drop-off Rate`   = dropoff
  ) |>
  mutate(across(c(`Form Start Rate`, `Drop-off Rate`),
                ~percent(., accuracy = 0.1))) |>
  kbl(align = c("l","c","c","c","c")) |>
  kable_styling(bootstrap_options = c("striped", "condensed", "bordered"),
                full_width = TRUE, font_size = 13) |>
  row_spec(0, bold = TRUE, background = "#7e82ad", color = "white")
```

::: callout-box
Drop-off exceeds **83% in every channel** — confirming a structural on-site barrier, not a channel-specific issue
:::

------------------------------------------------------------------------

## Strategic Recommendations {.section-slide}

::::::: rec-list
::: rec-item
[1]{.rec-num} **Redesign the Mid-Funnel Experience** — Strengthen CTAs, sharpen the value proposition, and simplify the path to the booking form. Highest-leverage fix without additional spend.
:::

::: rec-item
[2]{.rec-num} **Invest in Organic Search & SEO** — Converts at 2.3%, more than 2× the site average. Shift content and budget toward this channel.
:::

::: rec-item
[3]{.rec-num} **Audit UTM Tagging for Unassigned Traffic** — The site's highest-converting source (3.1%) is currently unlabeled. Classify it through proper UTM configuration and GA4 audit.
:::

::: rec-item
[4]{.rec-num} **Deploy Purchase Event Tracking via GTM** — Track completed purchases, not just bookings, to connect marketing to actual revenue outcomes.
:::
:::::::
