---
title: "Website Behavior & Booking Funnel Analysis"
subtitle: "Analytics Objective 5 | Google Analytics 4 | Jan 1 – Apr 21, 2026"
author: "Kathleen Juarez"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-title: "On this page"
    smooth-scroll: true
    theme: flatly
    css: ao5-styles.css
    page-layout: full
    embed-resources: true
    code-fold: true
    code-summary: "Show figure code"
execute:
  warning: false
  message: false
  echo: false
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(scales)
library(knitr)
library(kableExtra)

# ── Data Loading ──────────────────────────────────────────────────────────────

sessions_raw <- read_csv("A05_data_sessions.csv",
                         skip = 9, show_col_types = FALSE) |>
  rename(channel = 1, sessions = 2, total_users = 3) |>
  filter(!is.na(channel), channel != "")

events_raw <- read_csv("A05_data_ga4-events.csv",
                       skip = 9, show_col_types = FALSE) |>
  rename(event_name = 1, channel = 2, event_count = 3, total_users = 4) |>
  filter(!is.na(event_name), event_name != "")

# ── Funnel Stage Extraction ───────────────────────────────────────────────────

extract_event <- function(event_nm, col_name) {
  events_raw |>
    filter(event_name == event_nm) |>
    select(channel, {{ col_name }} := total_users)
}

page_view        <- extract_event("page_view",            page_views)
form_start       <- extract_event("form_start",           form_starts)
booking_complete <- extract_event("booking_complete",     bookings)
schedule_appt    <- extract_event("Schedule_Appointment", sched_appt_users)
contact_us       <- extract_event("contact_us",           contact_us_users)

# ── Build Analysis Table ──────────────────────────────────────────────────────

ga4_data <- sessions_raw |>
  select(channel, sessions) |>
  left_join(page_view,        by = "channel") |>
  left_join(form_start,       by = "channel") |>
  left_join(booking_complete, by = "channel") |>
  left_join(schedule_appt,    by = "channel") |>
  left_join(contact_us,       by = "channel") |>
  mutate(across(-channel, ~replace_na(., 0))) |>
  mutate(
    page_view_rate    = if_else(sessions > 0,    page_views  / sessions,   0),
    form_start_rate   = if_else(page_views > 0,  form_starts / page_views, 0),
    booking_rate      = if_else(form_starts > 0, bookings    / form_starts, 0),
    overall_conv_rate = if_else(sessions > 0,    bookings    / sessions,   0)
  )

totals <- ga4_data |>
  summarise(
    channel           = "TOTAL",
    sessions          = sum(sessions),
    page_views        = sum(page_views),
    form_starts       = sum(form_starts),
    bookings          = sum(bookings),
    sched_appt_users  = sum(sched_appt_users),
    contact_us_users  = sum(contact_us_users),
    overall_conv_rate = sum(bookings) / sum(sessions)
  )

site_avg <- totals$overall_conv_rate

# ── Shared ggplot theme ───────────────────────────────────────────────────────

theme_ao5 <- theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.y = element_line(color = "#edf0f7"),
    axis.ticks         = element_blank(),
    legend.position    = "none",
    plot.title         = element_text(face = "bold", size = 12, color = "#1a1a2e"),
    plot.subtitle      = element_text(size = 9.5, color = "#6b7280",
                                      margin = margin(b = 8))
  )
```

# 

# 

```{=html}
<div class="hero-banner">
  <h1 class="hero-title">Website Behavior &amp; Booking Funnel Analysis</h1>
  <p class="hero-subtitle">
    Using Google Analytics 4 event-level data to identify where users drop off,
    diagnose friction in the booking pathway, and evaluate acquisition channel
    quality across a 112-day observation window.
  </p>
  <div class="hero-stats">
    <div class="stat-chip">📈 GA4 Event Data</div>
     <div class="stat-chip">🔢 R / tidyverse</div>
    <div class="stat-chip">📉 Descriptive Funnel Analysis</div>
  </div>
</div>
```

## Project Overview

::: project-overview-intro
This report is part of a larger consulting engagement examining how digital marketing activity translates into customer bookings for a small service-based business. This objective focuses on website behavior and booking funnel optimization. Despite traffic arriving through multiple acquisition channels, the business lacked visibility into where visitors were dropping off and which channels were actually converting. This analysis uses GA4 event-level data to diagnose that gap.
:::

```{=html}
<div class="card-row">
  <div class="info-card problem-card">
    <div class="card-icon">⚠️</div>
    <h3>The Problem</h3>
    <ul>
      <li>No visibility into where users were leaving the site</li>
      <li>Traffic volume treated as a proxy for success</li>
      <li>Booking form drop-off untracked and unaddressed</li>
      <li>Channel quality vs. volume not distinguished</li>
    </ul>
  </div>
  <div class="info-card solution-card">
    <div class="card-icon">✅</div>
    <h3>The Analytical Goal</h3>
    <ul>
      <li>Map the four-stage booking funnel using GA4 events</li>
      <li>Identify the largest friction point in the user journey</li>
      <li>Segment funnel performance by traffic channel</li>
      <li>Validate patterns through subgroup analysis</li>
    </ul>
  </div>
</div>
```

------------------------------------------------------------------------

## Methodology

The method selected is a **Descriptive Funnel Analysis** using GA4 event-level exports. The funnel maps user progression through four sequential behavioral stages, using *total users* (not event count) throughout to avoid inflation from repeat-event firing within a single session.

### Funnel Stages

```{=html}
<div class="pipeline-steps">
  <div class="pipeline-step">
    <div class="step-content">
      <h4>Sessions</h4>
      <p>Entry point — all website visits; denominator for overall conversion rate</p>
    </div>
  </div>
  <div class="pipeline-arrow">→</div>
  <div class="pipeline-step">
    <div class="step-content">
      <h4>Page Views</h4>
      <p>Users engaging with site content; surface-level behavioral signal</p>
    </div>
  </div>
  <div class="pipeline-arrow">→</div>
  <div class="pipeline-step">
    <div class="step-content">
      <h4>Form Start</h4>
      <p>User begins filling the booking form — highest-intent observable action</p>
    </div>
  </div>
  <div class="pipeline-arrow">→</div>
  <div class="pipeline-step">
    <div class="step-content">
      <h4>Booking Complete</h4>
      <p>Form submitted — confirmed booking intent; the conversion event</p>
    </div>
  </div>
</div>
```

### Derived Metrics

| Metric             | Formula                             |
|--------------------|-------------------------------------|
| Page View Rate     | page_view users ÷ sessions          |
| Form Start Rate    | form_start users ÷ page_view users  |
| Booking Rate       | booking_complete ÷ form_start users |
| Overall Conv. Rate | booking_complete ÷ sessions         |

Traffic channel (GA4 Default Channel Group) is used as the primary segmentation variable, enabling comparison across six acquisition sources: Direct, Organic Search, Organic Social, Referral, Unassigned, and Paid Search.

------------------------------------------------------------------------

## Statistical Summary

```{r}
#| label: summary-stats
#| echo: false
#| code-fold: false

summary_df <- tibble(
  Metric = c(
    "Total Sessions",
    "Page View Users",
    "Form Start Users",
    "Bookings Completed",
    "Page View Rate",
    "Form Start Rate (of page views)",
    "Booking Rate (of form starts)",
    "Overall Conversion Rate",
    "Largest Drop-off Stage"
  ),
  Value = c(
    comma(totals$sessions),
    comma(totals$page_views),
    comma(totals$form_starts),
    comma(totals$bookings),
    percent(totals$page_views  / totals$sessions,   accuracy = 0.1),
    percent(totals$form_starts / totals$page_views, accuracy = 0.1),
    percent(totals$bookings    / totals$form_starts, accuracy = 0.1),
    percent(site_avg,                               accuracy = 0.1),
    paste0("page_view → form_start (",
           percent(1 - totals$form_starts / totals$page_views, accuracy = 0.1),
           " drop-off)")
  )
)

summary_df |>
  kbl(align = c("l", "r"), col.names = c("Metric", "Value")) |>
  kable_styling(
    bootstrap_options = c("striped", "condensed", "bordered"),
    full_width = FALSE,
    font_size  = 13
  ) |>
  row_spec(0, bold = TRUE, background = "#7e82ad", color = "white") |>
  row_spec(9, bold = TRUE)
```

------------------------------------------------------------------------

## Key Findings

The funnel narrows sharply and unevenly. The dominant friction point is the transition from page viewing to form initiation — a **91.9% drop-off rate** that holds across every traffic channel.

### Funnel: Most Users Exit Before Starting the Booking Form

```{r}
#| label: fig-funnel
#| fig-width: 7
#| fig-height: 4
#| fig-cap: "Unique users at each stage of the booking funnel. Red bar indicates the largest drop-off point."
#| echo: true
#| code-fold: true
#| code-summary: "Show figure code"

funnel_long <- tibble(
  stage = factor(
    c("Sessions", "Page Views", "Form Start", "Booking\nComplete"),
    levels = c("Sessions", "Page Views", "Form Start", "Booking\nComplete")
  ),
  count     = c(totals$sessions, totals$page_views,
                totals$form_starts, totals$bookings),
  highlight = c(FALSE, FALSE, TRUE, FALSE)
)

ggplot(funnel_long, aes(x = stage, y = count, fill = highlight)) +
  geom_col(width = 0.52) +
  geom_text(aes(label = comma(count)),
            vjust = -0.4, size = 3.8, fontface = "bold", color = "#1A1A1A") +
  scale_fill_manual(values = c("FALSE" = "#66ada9", "TRUE" = "#C0392B")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Most Users Exit Before Completing Any High-Intent Action",
       x = NULL, y = "Users") +
  theme_ao5
```

------------------------------------------------------------------------

### Conversion Rate by Traffic Channel

```{r}
#| label: fig-channels
#| fig-width: 7
#| fig-height: 4
#| fig-cap: "Booking conversion rate by traffic channel. Dashed line = site average."
#| echo: true
#| code-fold: true
#| code-summary: "Show figure code"

ga4_data |>
  mutate(
    channel   = fct_reorder(channel, overall_conv_rate),
    above_avg = overall_conv_rate > site_avg
  ) |>
  ggplot(aes(x = channel, y = overall_conv_rate, fill = above_avg)) +
  geom_col(width = 0.52) +
  geom_text(aes(label = percent(overall_conv_rate, accuracy = 0.1)),
            hjust = -0.15, size = 3.8, fontface = "bold", color = "#1A1A1A") +
  geom_hline(yintercept = site_avg, linetype = "dashed",
             color = "#888888", linewidth = 0.5) +
  annotate("text", x = 0.65, y = site_avg + 0.0025,
           label = paste0("Site avg: ", percent(site_avg, accuracy = 0.1)),
           size = 3.2, color = "#888888", hjust = 0) +
  coord_flip() +
  scale_fill_manual(values = c("FALSE" = "#7aa496", "TRUE" = "#66ada9")) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.28))) +
  labs(x = NULL, y = NULL) +
  theme_ao5 +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(color = "#edf0f7"))
```

------------------------------------------------------------------------

## Booking Funnel Performance by Traffic Channel

```{r}
#| label: tbl-funnel
#| echo: false
#| code-fold: false

ga4_data |>
  arrange(desc(overall_conv_rate)) |>
  bind_rows(totals) |>
  mutate(overall_conv_rate = percent(overall_conv_rate, accuracy = 0.1)) |>
  select(
    `Traffic Channel` = channel,
    Sessions          = sessions,
    `Page Views`      = page_views,
    `Contact Us`      = contact_us_users,
    `Form Starts`     = form_starts,
    `Sched. Appts`    = sched_appt_users,
    `Conv. Rate`      = overall_conv_rate
  ) |>
  kbl(align = c("l","c","c","c","c","c","c")) |>
  kable_styling(
    bootstrap_options = c("condensed", "bordered"),
    full_width = FALSE,
    font_size  = 13
  ) |>
  row_spec(0, bold = TRUE, background = "#7e82ad", color = "white") |>
  row_spec(seq(1, nrow(ga4_data), by = 2), background = "white",   color = "#333") |>
  row_spec(seq(2, nrow(ga4_data), by = 2), background = "#f3f5f9", color = "#333") |>
  row_spec(nrow(ga4_data) + 1, bold = TRUE,
           background = "#e8edf6", color = "#1a1a2e")
```

------------------------------------------------------------------------

### High-Intent Behavioral Signals

Schedule_Appointment and contact_us are link clicks to the page hosting the booking form. They are reported here as behavioral context rather than sequential funnel stages.

```{r}
#| label: tbl-signals
#| echo: false
#| code-fold: false

ga4_data |>
  arrange(desc(sessions)) |>
  bind_rows(totals) |>
  select(
    `Traffic Channel`    = channel,
    Sessions             = sessions,
    `Sched. Appt Clicks` = sched_appt_users,
    `Contact Us Clicks`  = contact_us_users,
    `Form Starts`        = form_starts,
    Bookings             = bookings
  ) |>
  kbl(align = c("l","c","c","c","c","c")) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "bordered"),
    full_width = FALSE,
    font_size  = 13
  ) |>
  row_spec(0, bold = TRUE, background = "#7e82ad", color = "white") |>
  row_spec(nrow(ga4_data) + 1, bold = TRUE, background = "#d6e8f5")
```

------------------------------------------------------------------------

## Recommendations

```{=html}
<div class="rec-list">
  <div class="rec-card">
    <div class="rec-body">
      <h4>Redesign the Mid-Funnel Experience</h4>
      <p>The 91.9% drop-off between page views and form starts is the highest-leverage fix available. Strengthening calls to action, sharpening the value proposition, and reducing booking form friction can increase appointments without requiring additional marketing spend.</p>
    </div>
  </div>
  <div class="rec-card">
    <div class="rec-body">
      <h4>Prioritize Organic Search &amp; SEO Investment</h4>
      <p>Organic Search converts at 2.3% — more than twice the site average and more than 4× the efficiency of Organic Social. Shifting content and budget investment toward this channel is likely to yield stronger returns than broad social reach alone.</p>
    </div>
  </div>
  <div class="rec-card">
    <div class="rec-body">
      <h4>Audit Channel Tracking &amp; UTM Configuration</h4>
      <p>Unassigned traffic converts at 3.1% — the highest rate on the site — but its source is currently invisible. Auditing UTM tagging and GA4 configuration to properly classify this traffic is a critical step toward understanding what is driving the best results.</p>
    </div>
  </div>
  <div class="rec-card">
    <div class="rec-body">
      <h4>Deploy Purchase Event Tracking via Google Tag Manager</h4>
      <p>Current tracking measures booking form submissions — not completed purchases. Configuring GTM to capture purchase completions will allow future analyses to connect marketing activity to actual revenue outcomes, enabling true ROI measurement.</p>
    </div>
  </div>
</div>
```

------------------------------------------------------------------------

# RevealJs Presentation

```{=html}
<iframe
  src="ao5-presentation.html"
  width="100%"
  height="520px"
  style="border: 1px solid #c5cfe0; border-radius: 8px; display: block;">
</iframe
```

## Appendix

### R Packages

| Package | Role |
|----|----|
| tidyverse | Data wrangling and visualization (dplyr, ggplot2, readr) |
| scales | Percent and comma formatting for labels |
| knitr / kableExtra | Table rendering with custom styling |
